Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data

### 2019-03-17 article/01_Introduction.Rmd
referenced the relevant work of Capuzzo et al.

How can major misconceptions like this arise and persist in European archaeology? Beyond missing data due to preservation and research history, another reason might be a lack of unified data repositories and universally acknowledged methods to measure the spatial and temporal distribution of cultural phenomena. Instead of directly querying archives with quantitative information, the typical process for the creation of large-scale syntheses is still to manually excerpt the literature of all relevant, local research traditions. The present article, in contrast, explores the former strategy. The first major question therefore is: *Can the main trends in the distribution of burial rites in Bronze Age Europe be detected in openly available bulk radiocarbon data?* This approach could set a precedent for a more reproducible and statistically sound research practice. If we could answer this first question with *yes*, then the quantitative comparison of macro regional trends could pave the way for an analysis of interaction intensity on a long-term continental scale: *Do time series constructed from radiocarbon data enable meaningful observations concerning cultural and spatial distance?*

>>>>

How can major misconceptions like this arise and persist in European archaeology? Beyond missing data due to preservation and research history, another reason might be a lack of unified data repositories and universally acknowledged methods to measure the spatial and temporal distribution of cultural phenomena. Instead of directly querying archives with quantitative information, the typical process for the creation of large-scale syntheses is still to manually excerpt the literature of all relevant, local research traditions. The present article, in contrast, explores the former strategy. The first major question therefore is: *Can the main trends in the distribution of burial rites in Bronze Age Europe be detected in openly available bulk radiocarbon data?*

The diffusion of the cremation burial tradition in the middle and late Bronze Age with radiocarbon data was already the subject of promising research in Central and South-Western Europe with focus on Catalonia and Switzerland [@capuzzoCulturalChangesSecond2015; @capuzzoInhumacionCremacionNordeste2017; @capuzzoSpacetemporalAnalysisRadiocarbon2014]. This approach based on the dedicated radiocarbon database EUBAR [@capuzzoEUBARDatabase14C2013] sets a precedent for a more reproducible and statistically sound research practice concerning diffusion of innovation in Bronze Age archaeology.

If radiocarbon data allows to reconstruct the diffusion of burial traditions, then the quantitative comparison of macro regional trends could also pave the way for an analysis of interaction intensity on a long-term continental scale. The second main question: *Do time series constructed from radiocarbon data enable meaningful observations concerning cultural and spatial distance?*
